---
title: "Excercise 3: Data Visualization"
author:
  name: Piyayut Chitchumnong
  affiliation: "[National Agricultural Big Data Center](https://nabc.go.th)"
format: 
  html:
    toc: true
    toc-location: left
    fig-width: 10
    fig-height: 6
---

## Load packages
```{r}
#| warning: false
library(tidyverse)
library(scales)
library(treemapify)
library(rnaturalearth)
library(sf)
library(viridis)
```

## Import data

- Import `baci_extra_i`  which is export by product (HS 2 digit) of exporting countries from 2011-2020. You have two options to load data.

**Option 1**: Import from internet.
```{r}
#| eval: false
url_baci_i <- "https://github.com/piyayut-ch/Workshop_URTA_2022/raw/main/data/baci_extra_i.rds"
baci_extra_i <- readRDS(gzcon(url(url_baci_i)))
# gzcon and url function are used to download data from github
```

**Option 2**: Import from your machine, if you open with the R project and have the file in folder data.
```{r}
path_baci <- here::here("data/baci_extra_i.rds")
baci_extra_i <- readRDS(path_baci)
```

```{r}
baci_extra_i
```

- Import `world` map using `rnaturalearth` package.

```{r}
#| column: screen-right
world <- ne_countries(scale = "small", returnclass = "sf")
plot(world['admin'])
```



## Question 1: Bar plot
Make top 20 exporters of `Meat` in 2018. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_bar <- baci_extra_i |>
  filter(t == ... & k == ...) |>
  mutate(i_name = fct_reorder(i_name, v)) |>
  arrange(desc(v)) |>
  head(20)

ggplot(data_bar, aes(x = ..., y = ..., fill = ...)) +
  geom_col() +
  scale_x_continuous(labels = comma)
```

### Solution
```{r}
#| column: screen-right
data_bar <- baci_extra_i |>
  filter(t == 2018 & k == "02") |>
  mutate(i_name = fct_reorder(i_name, v)) |> #reorder country name using factor
  arrange(desc(v)) |>
  head(20)

ggplot(data_bar, aes(x = v, y = i_name, fill = i_region)) +
  geom_col() +
  scale_x_continuous(labels = comma)
```
:::


## Question 2: Line plot
Make a line plot of `Flours, starches and malts` export of Thailand, Viet Nam and China from 2011 - 2020. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_line <- baci_extra_i |>
  filter(i_iso3c %in% c(...) & k == ...)

ggplot(data_line, aes(x = ..., y = ..., color = ...)) +
  geom_line() +
  scale_x_continuous(breaks = 2011:2020) +
  scale_y_continuous(labels = comma)
```

### Solution
```{r}
#| column: screen-right
data_line <- baci_extra_i |>
  filter(i_iso3c %in% c("THA", "VNM", "CHN") & k == "11")

ggplot(data_line, aes(x = t, y = v, color = i_name)) +
  geom_line() +
  scale_x_continuous(breaks = 2011:2020) +
  scale_y_continuous(labels = comma)
```
:::


## Question 3: Area plot
Make an area plot that shows export structure of India from 2011-2020 using share. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_area <- baci_extra_i |>
  filter(i_iso3c == ...) |>
  group_by(...) |>
  summarize(v = sum(v, na.rm = TRUE)) |>
  ungroup() |>
  group_by(t) |>
  mutate(pct = ...)

ggplot(data_area, aes(x = ..., y = ..., fill = ...)) +
  geom_area(alpha = 0.8, size = 0.5, color = "black") +
  scale_x_continuous(breaks = 2011:2020) +
  scale_y_continuous(labels = comma)
```

### Solution
```{r}
#| column: screen-right
#| warning: false
data_area <- baci_extra_i |>
  filter(i_iso3c == "IDN") |>
  group_by(t, k_sector) |>
  summarize(v = sum(v, na.rm = TRUE)) |>
  ungroup() |>
  group_by(t) |>
  mutate(pct = v / sum(v))

ggplot(data_area, aes(x = t, y = pct, fill = k_sector)) +
  geom_area(alpha = 0.8, size = 0.5, color = "black") +
  scale_x_continuous(breaks = 2011:2020) +
  scale_y_continuous(labels = comma)
```
:::


## Question 4: Histogram
Make a histogram of distribution of export of trade value of `Oil seeds and oleaginous fruits` in 2020 in log. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_hist <- baci_extra |>
  filter(t == ..., k == ...)

ggplot(data_hist, aes(x = ...)) +
  geom_histogram(
    bins = 30,
    fill = "navyblue",
    color = "white"
  )
```

### Solution
```{r}
#| column: screen-right
data_hist <- baci_extra_i |>
  filter(t == 2020, k == "12")

ggplot(data_hist, aes(x = log(v))) +
  geom_histogram(
    bins = 30,
    fill = "navyblue",
    color = "white"
  )
```
:::


## Question 5: Box plot
Make a box plot of distribution of log of trade value in 2020 of

- `Meat`
- `Fish`
- `Cereals`
- `Flours, starches and malts`
- `Oil seeds and oleaginous fruits`

Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_boxplot <- baci_extra_i |>
  filter(k %in% c(...), t == ...)

ggplot(data_boxplot, aes(x = ..., y = log(...), fill = ...)) + 
  geom_boxplot() +
  coord_flip()
```

### Solution
```{r}
#| column: screen-right
data_boxplot <- baci_extra_i |>
  filter(k %in% c("02", "03", "10", "11", "12"), t == 2020)

ggplot(data_boxplot, aes(x = k_name, y = log(v), fill = k_name)) + 
  geom_boxplot() +
  coord_flip()
```
:::


## Question 6: Treemap
Make a treemap plot of Cambodia in 2011. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_treemap <- baci_extra_i |>
  filter(t == ..., i_iso3c == ...)

ggplot(
  data_treemap,
  aes(area = ...,
      fill = ...,
      subgroup = ...,
      label = ...)
  ) +
  geom_treemap(color = "white", start = "topleft") +
  geom_treemap_subgroup_border(color = "grey40", start = "topleft") +
  geom_treemap_text(
    color = "grey30",
    place = "topleft",
    start = "topleft",
    reflow = TRUE) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1))
```

### Solution
```{r}
#| column: screen-right
data_treemap <- baci_extra_i |>
  filter(t == 2011, i_iso3c == "KHM")

ggplot(
  data_treemap,
  aes(area = v,
      fill = k_sector,
      subgroup = k_sector,
      label = k_name)
  ) +
  geom_treemap(color = "white", start = "topleft") +
  geom_treemap_subgroup_border(color = "grey40", start = "topleft") +
  geom_treemap_text(
    color = "grey30",
    place = "topleft",
    start = "topleft",
    reflow = TRUE) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1))
```
:::


## Question 7: Scatter plot
Make a scatter plot of unit price (x) and export value (y) of `Oil seeds and oleaginous fruits` in 2015. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
data_scatter <- baci_extra_i |>
  filter(t == ..., k == ...) |>
  mutate(
    unit_price = v * 1e6 / q,
    gdp = gdp / 1e6
  )

ggplot(
  data_scatter,
  aes(x = ...,
      y = ...,
      size = ...,
      color = ...)
  ) +
  geom_point(alpha = 0.8) +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  scale_size(range = c(1, 20), guide = NULL)
```

### Solution
```{r}
#| column: screen-right
data_scatter <- baci_extra_i |>
  filter(t == 2015, k == "12") |>
  mutate(
    unit_price = v * 1e6 / q,
    gdp = gdp / 1e6
  )

ggplot(
  data_scatter,
  aes(x = v,
      y = unit_price,
      size = gdp,
      color = i_region)
  ) +
  geom_point(alpha = 0.8) +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  scale_size(range = c(1, 20), guide = NULL)
```
:::


## Question 8: Map
Make a cholorpleth map (heatmap) of `Diary products` in year 2019. Fill `...` in the code snippet.

::: {.panel-tabset}
### Code
```{r}
#| eval: false
mytheme <- theme(
  panel.background = element_rect(fill = 'black'),
  panel.grid.major = element_line(color = 'black')
)

data_map <- baci_extra_i |> filter(t == ..., k == ...)

world_baci <- world |> 
  filter(continent != "Antarctica") |>
  left_join(
    data_map,
    by = c("adm0_a3" = "i_iso3c")
  )

ggplot(world_baci, aes(fill = ...)) +
  geom_sf(color = "grey60") +
  scale_fill_viridis(option = "magma", begin = 0.1, end = 1) +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_colorbar(
      title = "Export Value",
      barwidth = 30,
      direction = "horizontal"
    )
  ) +
  mytheme
```

### Solution
```{r}
#| column: screen-right
mytheme <- theme(
  panel.background = element_rect(fill = 'black'),
  panel.grid.major = element_line(color = 'black')
)

data_map <- baci_extra_i |> filter(t == 2019, k == "04")

world_baci <- world |> 
  filter(continent != "Antarctica") |>
  left_join(
    data_map,
    by = c("adm0_a3" = "i_iso3c")
  )

ggplot(world_baci, aes(fill = v)) +
  geom_sf(color = "grey60") +
  scale_fill_viridis(option = "magma", begin = 0.1, end = 1) +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_colorbar(
      title = "Export Value",
      barwidth = 30,
      direction = "horizontal"
    )
  ) +
  mytheme
```
:::